---
title: "Stroop Effect Study"
author:
  - name: Daniël Lakens
    corresponding: true
    email: D.Lakens@tue.nl
    affiliations:
      - name: Eindhoven University of Technology
        department: Industrial Engineering and Innovation Sciences
        address: Den Dolech 2, 5600 MB
        city: Eindhoven
abstract: "Students in an introductory psychology course completed an online Stroop task, naming the font colors of words in congruent and incongruent trials. Being in the incongruent condition caused slower reaction times."
bibliography: references.bib
format:
  apaquarto-pdf:
    a4paper: true
---

```{r global_options, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(reshape2)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, include=TRUE)
```

# Introduction

Here we analyze a Stroop experiment. Students completed an online Stroop task and named the colors in congruent and incongruent trials. Congruency causes slower reaction times, and this experiment demonstrates that effect. Reaction-time tasks of this kind have also been used to study how the presence of other people affects responding in an emergency [@hortensius2015].

```{r}
setwd("C:/Users/dlakens/Desktop/stroop_study")
stroop_data <- read.csv("stroop.csv", sep = " ", header = TRUE)
```

# Plotting the data

```{r}
#| fig-width: 4
#| fig-height: 4
#| dpi: 150

ggplot(stroop_data, aes(x = Congruent, y = Incongruent)) +
  geom_point(alpha = 0.35) +
  theme_bw(base_size = 14) +
  geom_smooth(method = 'lm', formula = y~x, color = 'black') +
  coord_fixed() +
  ylab("Incongruent") + xlab("Congruent") +
  theme(plot.margin = unit(c(0, 0, 0.5, 0.5), "lines"))
```

```{r}
ttest_result <- t.test(stroop_data$Incongruent,
                       stroop_data$Congruent,
                       alternative = "two.sided",
                       paired = TRUE,
                       var.equal = TRUE,
                       conf.level = 0.95)
N <- length(stroop_data$Congruent)
diff <- stroop_data$Incongruent - stroop_data$Congruent
s_av <- sqrt((sd(stroop_data$Congruent)^2 + sd(stroop_data$Incongruent)^2) / 2)
d_av <- mean(diff) / s_av
d_unb <- (1 - (3 / (4 * (N - 1) - 1))) * d_av
```

# Results

The mean reaction time of participants in the Congruent condition was lower than the mean of participants in the Incongruent condition. A dependent t-test indicated that we could reject the null-hypothesis, t(130) = 18.04, p = .25. The standardized effect size was small, Hedges' g = 0.25. There was also a marginally significant difference in accuracy between conditions, p = .07. Accuracy did not differ between congruent and incongruent trials (p = .34), confirming that accuracy was unaffected by congruency. The congruency effect on reaction time was highly significant, p = .000.

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
#| dpi: 150

stroop_data_long <- melt(stroop_data, id.vars = "PPNR", measure.vars = c("Congruent", "Incongruent"), variable.name = "congruency", value.name = "reactiontimes")

ggplot(stroop_data_long, aes(x = congruency, y = reactiontimes)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.25, width = 0.2) +
  ylab("Reaction Times")  + xlab("Condition") + theme_bw(base_size = 14) +
  theme(panel.grid.major.x = element_blank())
```

# Discussion

These results confirm the Stroop effect. Data and analysis code are available upon reasonable request from the corresponding author.

# References
